Challenges and Solutions of Developing and Implementing a Novel Desktop-as-a-Service

### 2024-05-01 paper.tex
Add figures to chapters performance and stability

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Novel Desktop-as-a-Service}
%
\titlerunning{Challenges and Solutions of Developing and
	Implementing a Novel DaaS}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun \and
	Johannes Bouché}
% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
% 	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
	Frankfurt University of Applied Sciences,\\
	Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	Desktop-as-a-Service (DaaS) is a cloud service category with a tremendous potential for improving the everyday work of running and using applications. However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past. Our novel DaaS solution called DESIGN not only overcomes the limitations of existing solutions but also offers significant benefits. It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side. This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.

@@

As a result of these considerations, we aim to support
any Windows or Linux-based application in virtual machines
and support any Linux- and Wine-based applications for container environments.

@@

We utilize the Guacamole desktop gateway implementation,
supporting widely-used frame-buffer protocols like VNC and RDP,
along with relevant tuning options.
Guacamole also enables audio support and integration with external devices
through a JavaScript-based web client.

However, integrating Guacamole into the system
requires consideration of additional factors not handled by the software.
Authentication mechanisms for example are particularly important in a DaaS context,
as users expect seamless authentication across the whole system
without multiple logins per session.
Additionally, safeguarding sensitive information like user passwords is crucial
for maintaining system security and user trust.
To mitigate the risk of unintended system manipulation,
Guacamole was integrated with a proxied approach and by using websockets.
This ensures that credentials of a generated instance remain secret to the host system
while still providing authorized, passwordless access to the hosted application.

@@

@@

From our observations, components requiring realtime and low-latency performance,
should physically be kept as close to the core hosting system as possible.
This includes all data and communication relevant to
audio, video, mouse, and keyboard transmission,
as well as all backend services for hosting containers and virtual machines.
Ideally, these components should be hosted natively on the host system
for virtual network and hardware access
or at least be connected to the same physical network
in order to minimize network latencies.

@@

@@

Vertical scaling of DaaS nodes is not currently planned,
but administrators are free to use custom third-party toolchains,
such as GPU virtualization, for vertical scaling on a per-host basis.
Although vertical scaling within a host is beyond the scope of our research project,
scaling the instances hosted within the system is feasible for both dimensions.

For vertical scaling of instances, we provide appropriate parameters
and API endpoints to manipulate resource assignments,
allowing for arbitrary configurations.
For horizontal scaling of instances, the built-in mechanisms
of each backend service are automatically configured
and utilized to distribute instances across a pool of connected nodes.
For virtual machines, Proxmox implements such mechanisms
as part of its 'Proxmox VE High Availability (HA)' cluster.
For containers, similar concepts can be integrated with any backend service,
such as Kubernetes.
Within these clusters,
DaaS instances are entirely agnostic to their hosting system
and only need to be reachable via a known hostname or IP as a single requirement.

\subsubsection{Redundancy of Components}

Redundancy is crucial in distributed systems,
particularly in DaaS architectures,
to safeguard against service interruptions and data loss
by duplicating essential components, functions, and data.
This design principle ensures high availability and reliability,
even during component failures.
In the following, we provide a list of components eligible for redundancy,
briefly describing the most relevant strategies in our architecture.

\begin{itemize}
	\item \textbf{Frontend/Backend:}
	      In our final version,
	      redundancy of the frontend and backend components
	      will enable configuration for distributing multiple instances
	      to dedicated systems and utilizing standardized load-balancing strategies.
	      This setup allows the frontend or backend
	      to serve as a gateway between the user and the backend service
	      containing all relevant instances.
	      Both components contribute to redundancy
	      by distributing requests and responses to their respecitive recipients
	      instead of processing them on their own.
	\item \textbf{Database:}
	      The database is integrated as a backend service
	      and is therefore independent of surrounding components and contexts.
	      It can be distributed to a dedicated system if necessary.
	      Modern database solutions, like MariaDB,
	      offer rich features for maintaining specific data consistency strategies
	      such as data replication, clustering, failover mechanisms,
	      and RAID control.
	      Currently, redundancies in our database system are not planned
	      but can be easily applied by users using such database-specific features
	      and a modified database adapter.
	\item \textbf{Authentication: }
	      Authentication within our architecture is implemented as a backend service,
	      allowing for distribution to any dedicated system.
	      It can also be targeted by standardized load-balancing mechanisms
	      and remains fully agnostic of surrounding components
	      or calling contexts.
	      The authentication system maintains its credential registry
	      in a separate database, along with a customizable set
	      of user and group privileges.
	      Additionally, it implements the well-defined OAuth 2.0 standard
	      to facilitate remote authentication.
	      In our final version,
	      a customizable amount of redundant systems may be configured.
	\item \textbf{Distributed Filesystem:}
	      Within our architecture, we utilize the distributed filesystem Ceph,
	      which offers object, block, and file storage capabilities.
	      Redundancy in Ceph is implicitly achieved
	      through its self-healing and self-managing features, employing
	      the CRUSH algorithm (Controlled Replication Under Scalable Hashing).
	      Data availability and fault tolerance are ensured
	      by automatically replicating data across different devices.
	      Our architecture leverages these capabilities
	      to reliably provide file storage to host systems
	      and all hosted DaaS instances,
	      while also offering block storage to the infrastructural backend services.
	      Hosting DaaS instances directly from the distributed block storage
	      enables migration and redundancy of instances or in other words, horizontal scaling.
	\item \textbf{Instances:} Redundancy of instances
	      It is not a necessary condition for DaaS systems in general,
	      but applying such strategies under certain conditions might be plausible.
	      By using highly available clusters within our backend services
	      redundancy can be achieved by using the corresponding API endpoints
	      previously described and by parameterizing them appropriately.
\end{itemize}

@@

This ensures that any change triggered by a node within the DaaS domain
is immediately available to any other node relying in some way
on that particular information change.
Additionally, any extra information needed by a component must be fully contained
within the related request or response.
This enables arbitrarily distributed configurations as long as request
and response messages reach their desired target eventually
and as long relevant backend services remain available.

@@

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{positioning, shapes.geometric}
\usepackage{adjustbox}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Novel Desktop-as-a-Service}
%
\titlerunning{Challenges and Solutions of Developing and
	Implementing a Novel DaaS}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun \and
	Johannes Bouché}
% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
% 	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
	Frankfurt University of Applied Sciences,\\
	Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	Desktop-as-a-Service (DaaS) is a cloud service category with a tremendous potential for improving the everyday work of running and using applications. However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past. Our novel DaaS solution called DESIGN not only overcomes the limitations of existing solutions but also offers significant benefits. It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side. This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.

@@

Figure~\ref{hosting_scenarios} visualizes the available hosting startegies for our platform.
We aim to support any Windows or Linux-based application in virtual machines
and support any Linux- and Wine-based applications for container environments.

\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]
			%Type Column
			\node[minimum width=2cm] at (-4,-2) {IaaS};
			\node[minimum width=2cm] at (-4,-1) {PaaS};
			\node[minimum width=2cm] at (-4,0) {SaaS};
			\node[minimum width=2cm] at (-4,1) {DaaS};
			% DaaS
			\node[draw, minimum width=12cm, fill=gray!10] at (3,1) {DESIGN Frontend};
			%Linux
			\node[draw, minimum width=6cm, fill=gray!50] at (0,0) {Linux App};
			\node[draw, minimum width=6cm, fill=gray!80] at (0,-1) {Linux OS};
			\node[draw, minimum width=3cm, fill=gray!100] at (-1.5,-2) {Virtual Machine};
			\node[draw, minimum width=3cm, fill=gray!100] at (1.5,-2) {Container};
			%Windows
			\node[draw, minimum width=6cm, fill=gray!50] at (6,0) {Windows App};
			\node[draw, minimum width=3cm, fill=gray!80] at (4.5,-1) {Windows OS};
			\node[draw, minimum width=3cm, fill=gray!80] at (7.5,-1) {Linux OS, Wine};
			\node[draw, minimum width=3cm, fill=gray!100] at (4.5,-2) {Virtual Machine};
			\node[draw, minimum width=3cm, fill=gray!100] at (7.5,-2) {Container};
		\end{tikzpicture}
	\end{adjustbox}
	\caption{Application hosting strategies.}\label{hosting_scenarios}
\end{figure}

@@

We utilize the Guacamole desktop gateway implementation,
supporting widely-used frame-buffer protocols like VNC and RDP,
along with relevant tuning options.
Guacamole also enables audio support and integration with external devices
through a JavaScript-based web client (guacamole.js).
As depicted in figure~\ref{websocket_proxy},
the proxy component recives HTTP requests from the guacamole.js webclient
and forwarded by the backend by using regular WebSocket opcodes (Guacamole protocol).
Such communication is forwarded to the Guacamole daemon (guacd)
which in turn communicates to each instance
by using the frame-buffer protocol of choice (VNC or RDP).

However, integrating Guacamole into the system
requires consideration of additional factors besides the frame-buffer protocols itself.
Authentication mechanisms for example are particularly important in a DaaS context,
as users expect seamless authentication across the whole system
without multiple logins per session.
Additionally, safeguarding sensitive information like user passwords is crucial
for maintaining system security and user trust.
To mitigate the risk of unintended system manipulation, Guacamole was integrated
in a proxied approach and by using websockets.
This ensures that credentials of a generated instance remain secret to the host system
while still providing authorized, passwordless access to the hosted application.

@@

\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]

			% Nodes center row
			\node[draw, minimum width=3cm, fill=gray!10] (browser) at (0,0) {guacamole.js };
			\node[draw, minimum width=3cm, fill=gray!50] (back) at (4,0) {DESIGN Backend};
			\node[draw, minimum width=3cm, fill=gray!50] (proxy) at (8,0) {Proxy};
			\node[draw, minimum width=3cm, fill=gray!50] (guacd) at (12,0) {guacd};
			\node[draw, minimum width=3cm, fill=gray!100] (inst) at (16,0) {Instances};
			% Arrows center row
			\draw[<->, thick, black] (browser.east) -- (back.west);
			\draw[<->, thick, black] (back.east) -- (proxy.west);
			\draw[<->, thick, black] (proxy.east) -- (guacd.west);
			\draw[<->, thick, black] (guacd.east) -- (inst.west);
			% Upper row
			\node[draw, minimum width=3cm, fill=gray!10] (client-backend) at (2,2) {HTTP/WebSocket};
			\node[draw, minimum width=3cm, fill=gray!50] (backend-proxy) at (6,2) {HTTP/WebSocket};
			\node[draw, minimum width=3cm, fill=gray!50] (proxy-guacd) at (10,2) {Guacamole protocol};
			\node[draw, minimum width=3cm, fill=gray!100] (guacd-inst) at (14,2) {Framebuffer protocols/\\VNC,RDP};
			% Dashed lines upper row
			\draw[-, dashed, thick, black] (2,0) -- (2,1.5);
			\draw[-, dashed, thick, black] (6,0) -- (6,1.5);
			\draw[-, dashed, thick, black] (10,0) -- (10,1.5);
			\draw[-, dashed, thick, black] (14,0) -- (14,1.5);
			% Arrow bottom row
			\draw[<->, thick, black] (0,-1) -- (3.8,-1) node[midway, below, font=\footnotesize] {Client Communication};
			\draw[<->, thick, black] (4,-1) -- (11.8,-1)node[midway, below, font=\footnotesize] {Backend Communication};
			\draw[<->, thick, black] (12,-1) -- (16,-1)node[midway, below, font=\footnotesize] {Instance Communication};
		\end{tikzpicture}
	\end{adjustbox}
	\caption{WebSocket Proxy Components.}\label{websocket_proxy}
\end{figure}

@@

From our observations, components requiring realtime and low-latency performance,
should physically be kept as close to the core hosting system as possible
as shown in figure~\ref{perf_internal_communication}.
This includes all data and communication relevant to
audio, video, mouse, and keyboard transmission,
as well as all backend services for hosting containers and virtual machines.
Ideally, these components should be hosted natively on the host system
for virtual network and hardware access
or at least be connected to the same physical network
in order to minimize network latencies.

@@

\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]
			% Standalone Realtime
			\node[draw, minimum width=8cm, fill=gray!10] (singlefront) at (0,-0.5) {DESIGN Frontend};
			\node[draw, minimum width=8cm, fill=gray!30] (singleback) at (+0,-2) {DESIGN Backend};
			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-3) {Infrastructure\\(Proxmox/Docker) };
			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-4) {Proxy};
			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-5) {Backend API};
			% Standalone Services
			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-3) {Database};
			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-4) {Filesystem};
			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-5) {Authentication};
			%Standalone Caption
			\node[minimum width=8cm] (browser) at (0,-6) {Standalone};
			% Distributed Realtime
			\node[draw, minimum width=10cm, fill=gray!10] (distfront) at (10,-0.5) {DESIGN Frontend};
			\node[draw, minimum width=10cm, fill=gray!30] (distback) at (10,-2) {DESIGN Backend};
			\node[draw, minimum width=4cm, fill=gray!100] (inf) at (7,-3) {Infrastructure\\(Proxmox/Docker) };
			\node[draw, minimum width=4cm, fill=gray!100] (proxy) at (7,-4) {Proxy};
			\node[draw, minimum width=4cm, fill=gray!100] (api) at (7,-5) {Backend API};
			% Distributed Services
			\node[draw, minimum width=4cm, fill=gray!50] (db) at (13,-3) {Database};
			\node[draw, minimum width=4cm, fill=gray!50] (fs) at (13,-4) {Filesystem};
			\node[draw, minimum width=4cm, fill=gray!50] (auth) at (13,-5) {Authentication};
			% Network
			\node[draw, minimum width=1cm, minimum height=3cm, fill=white] (browser) at (10,-4) {N\\E\\T\\W\\O\\R\\K};
			% Distributed Caption
			\node[minimum width=10cm] (browser) at (10,-6) {Distributed};
			% Arrows back<->front
			\draw[<->, thick, black] (singleback.north) -- (singlefront.south);
			\draw[<->, thick, black] (distback.north) -- (distfront.south);
			% Arrows realtime
			\draw[<->, thick, black] (inf.east) -- ++(0.5,0);
			\draw[<->, thick, black] (proxy.east) -- ++(0.5,0);
			\draw[<->, thick, black] (api.east) -- ++(0.5,0);
			% Arrows services
			\draw[<->, thick, black] (db.west) ++(-0.5,0) -- ++(0.5,0);
			\draw[<->, thick, black] (fs.west) ++(-0.5,0) -- ++(0.5,0);
			\draw[<->, thick, black] (auth.west) ++(-0.5,0) -- ++(0.5,0);
		\end{tikzpicture}
	\end{adjustbox}
	\caption{Internal Communication Flow}\label{perf_internal_communication}
\end{figure}

@@

Vertical scaling of DaaS nodes is not currently planned,
but administrators are free to use custom third-party toolchains,
such as GPU virtualization, for vertical scaling on a per-host basis.
Although vertical scaling within a host
is beyond the scope of our research project,
scaling the instances hosted within the system is feasible for both dimensions.
For vertical scaling of instances, we provide appropriate parameters
and API endpoints to manipulate resource assignments,
allowing for arbitrary configurations.

For horizontal scaling of instances, the built-in mechanisms
of each backend service are automatically configured
and utilized to distribute instances across a pool of connected nodes.
For virtual machines, Proxmox implements such mechanisms
as part of its `Proxmox VE High Availability (HA)' cluster.
For containers, similar concepts can be integrated with any backend service,
such as Kubernetes.
Within these clusters,
DaaS instances are entirely agnostic to their hosting system
and only need to be reachable
via a known hostname or IP as a single requirement.
As shown in figure~\ref{scale_objects_horizontal},
we provide a globally maintained state of data, authentication and storage
which is accessible either by horizontally scaled backends and frontends
or otherwise,
by horizontally and vertically scaled instances in our DaaS object domain.

\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum width=4cm, minimum height=1cm, align=center}]

			% User
			\node[draw, minimum width=4cm] (user) at (5,1.5) {User};
			% Clusters
			\node[] at (-5,-0.5) {Frontend Cluster\\(Horizontal)};
			\node[] at (-5,-2.5) {Backend Cluster\\(Horizontal)};
			\node[] at (-5,-4.5) {Object Domain\\(Horizontal and Vertical)};

			% Node1
			\node[draw, fill=gray!10]  (front1) at (0,-0.5) {DESIGN Frontend 1};
			\node[draw, fill=gray!30]  (back1)  at (0,-2.0) {DESIGN Backend};
			\node[draw, fill=gray!50]  (real1)   at (0,-3.0) {Realtime Components};
			\node[draw, fill=gray!100] (inst1)  at (0,-4.5) {DaaS Instance (ID=1234)};
			% Node2
			\node[draw, fill=gray!10]  (front2) at (5,-0.5) {DESIGN Frontend};
			\node[draw, fill=gray!30]  (back2) at (5,-2.0) {DESIGN Backend};
			\node[draw, fill=gray!50]  (real2) at (5,-3.0) {Realtime Components};
			\node[draw, fill=gray!100] (inst2)  at (5,-4.5) {DaaS Instance (ID=1234)};
			% Node3
			\node[draw, fill=gray!10]  (front3) at (10,-0.5) {DESIGN Frontend};
			\node[draw, fill=gray!30]  (back3) at (10,-2.0) {DESIGN Backend};
			\node[draw, fill=gray!50]  (real3) at (10,-3.0) {Realtime Components};
			\node[draw, fill=gray!100] (inst3)  at (10,-4.5) {DaaS Instance (ID=1234)};
			% Domain
			\node[draw, dashed, minimum width=14.5cm, minimum height=1.25cm]  (domfront) at (5,-0.5){};
			\node[draw, dashed, minimum width=14.5cm, minimum height=2.25cm]  (domback) at (5,-2.5){};
			\node[draw, dashed, minimum width=14.5cm, minimum height=1.25cm]  (dominst) at (5,-4.5){};
			% Backend Services
			\node[draw, minimum width=2cm, minimum height=5cm] (backends) at (14,-2.5){Backend\\Services};
			% Arrows back<->Inst
			\draw[<->, thick, black] (inst1.north) -- (real1.south);
			\draw[<->, thick, black] (inst2.north) -- (real2.south);
			\draw[<->, thick, black] (inst3.north) -- (real3.south);
			% % Arrows back<->front
			\draw[<->, thick, black] (front1.south) -- (back1.north);
			\draw[<->, thick, black] (front2.south) -- (back2.north);
			\draw[<->, thick, black] (front3.south) -- (back3.north);
			% % Arrows Insts
			\draw[<->, thick, black] (inst1.east) -- (inst2.west);
			\draw[<->, thick, black] (inst2.east) -- (inst3.west);
			% % Arrows User
			\draw[->, thick, black] (user.south) -- (front1.north);
			\draw[->, thick, black] (user.south) -- (front2.north);
			\draw[->, thick, black] (user.south) -- (front3.north);
			% % Arrows Backend
			\draw[<->, thick, black] (domfront.east) -- ++(0.75,0);
			\draw[<->, thick, black] (domback.east) --  ++(0.75,0);
			\draw[<->, thick, black] (dominst.east) --  ++(0.75,0);
		\end{tikzpicture}
	\end{adjustbox}
	\caption{DESIGN Component Scaling}\label{scale_objects_horizontal}
\end{figure}

\subsubsection{Redundancy of Backend Services}

Redundancy is crucial in distributed systems,
particularly in DaaS architectures,
to safeguard against service interruptions and data loss
by duplicating essential components, functions, and data.
This design principle ensures high availability and reliability,
even during component failures.
Figure~\ref{service_redundancy} depicts all major DESIGN components eligible for redundancy
and which will be briefly described following:
\begin{itemize}
	\item \textbf{Frontend/Backend:}
	      In our final version,
	      redundancy of the frontend and backend components
	      will enable configuration for distributing multiple instances
	      to dedicated systems and utilizing standardized load-balancing strategies.
	      This setup allows the frontend or backend
	      to serve as a gateway between the user and the backend service
	      containing all relevant instances.
	      Both components contribute to redundancy
	      by distributing requests and responses to their respecitive recipients
	      instead of processing them on their own.
	\item \textbf{Instances:} Redundancy of instances
	      It is not a necessary condition for DaaS systems in general,
	      but applying such strategies under certain conditions might be plausible.
	      By using highly available clusters within our backend services
	      redundancy can be achieved by using the corresponding API endpoints
	      previously described and by parameterizing them appropriately.
	\item \textbf{Database:}
	      The database is integrated as a backend service
	      and is therefore independent of surrounding components and contexts.
	      It can be distributed to a dedicated system if necessary.
	      Modern database solutions, like MariaDB,
	      offer rich features for maintaining specific data consistency strategies
	      such as data replication, clustering, failover mechanisms,
	      and RAID control.
	      Currently, redundancies in our database system are not planned
	      but can be easily applied by users using such database-specific features
	      and a modified database adapter.
	\item \textbf{Authentication: }
	      Authentication within our architecture is implemented as a backend service,
	      allowing for distribution to any dedicated system.
	      It can also be targeted by standardized load-balancing mechanisms
	      and remains fully agnostic of surrounding components
	      or calling contexts.
	      The authentication system maintains its credential registry
	      in a separate database, along with a customizable set
	      of user and group privileges.
	      Additionally, it implements the well-defined OAuth 2.0 standard
	      to facilitate remote authentication.
	      In our final version,
	      a customizable amount of redundant systems may be configured.
	\item \textbf{Distributed Filesystem:}
	      Within our architecture, we utilize the distributed filesystem Ceph,
	      which offers object, block, and file storage capabilities.
	      Redundancy in Ceph is implicitly achieved
	      through its self-healing and self-managing features, employing
	      the CRUSH algorithm (Controlled Replication Under Scalable Hashing).
	      Data availability and fault tolerance are ensured
	      by automatically replicating data across different devices.
	      Our architecture leverages these capabilities
	      to reliably provide file storage to host systems
	      and all hosted DaaS instances,
	      while also offering block storage to the infrastructural backend services.
	      Hosting DaaS instances directly from the distributed block storage
	      enables migration and redundancy of instances or in other words, horizontal scaling.
\end{itemize}
\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum height=1.5cm, align=center}]
			% Node
			\node[draw, minimum width=4cm, fill=gray!10] (front0) at (-4,-3.0) {DESIGN Frontend};
			\node[draw, minimum width=4cm, fill=gray!30] (back0) at (-4,-4.5) {DESIGN Backend};
			\node[draw, minimum width=4cm, fill=gray!50] (inf0) at (-4,-6.0) {DESIGN Infrastructure};
			\node[draw, minimum width=4cm, fill=gray!10] (db0) at (0,-3.0) {Database};
			\node[draw, minimum width=4cm, fill=gray!30] (auth0) at (0,-4.5) {Authentication};
			\node[draw, minimum width=4cm, fill=gray!50] (fs0) at (0,-6.0) {Storage};
			% Frontends
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front1) at (-9,-2.5) {Frontend 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front2) at (-9,-3.0) {Frontend 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front3) at (-9,-3.5) {Frontend 3};
			\draw[->, dashed, thick, black] (front0.west) -- (front1.east);
			\draw[->, dashed, thick, black] (front0.west) -- (front2.east);
			\draw[->, dashed, thick, black] (front0.west) -- (front3.east);
			% Backends
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back1) at (-9,-4.0) {Backend 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back2) at (-9,-4.5) {Backend 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back3) at (-9,-5.0) {Backend 3};
			\draw[->, dashed, thick, black] (back0.west) -- (back1.east);
			\draw[->, dashed, thick, black] (back0.west) -- (back2.east);
			\draw[->, dashed, thick, black] (back0.west) -- (back3.east);
			% Instances
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst1) at (-9,-5.5) {Instances 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst2) at (-9,-6.0) {Instances 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst3) at (-9,-6.5) {Instances 3};
			\draw[->, dashed, thick, black] (inf0.west) -- (inst1.east);
			\draw[->, dashed, thick, black] (inf0.west) -- (inst2.east);
			\draw[->, dashed, thick, black] (inf0.west) -- (inst3.east);
			% DB
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db1) at (5,-2.5) {Database 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db2) at (5,-3.0) {Database 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db3) at (5,-3.5) {Database 3};
			\draw[->, dashed, thick, black] (db0.east) -- (db1.west);
			\draw[->, dashed, thick, black] (db0.east) -- (db2.west);
			\draw[->, dashed, thick, black] (db0.east) -- (db3.west);
			% Auth
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth1) at (5,-4.0) {Authentication 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth2) at (5,-4.5) {Authentication 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth3) at (5,-5.0) {Authentication 3};
			\draw[->, dashed, thick, black] (auth0.east) -- (auth1.west);
			\draw[->, dashed, thick, black] (auth0.east) -- (auth2.west);
			\draw[->, dashed, thick, black] (auth0.east) -- (auth3.west);
			% Fs
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs1) at (5,-5.5) {Storage 1};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs2) at (5,-6.0) {Storage 2};
			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs3) at (5,-6.5) {Storage 3};
			\draw[->, dashed, thick, black] (fs0.east) -- (fs1.west);
			\draw[->, dashed, thick, black] (fs0.east) -- (fs2.west);
			\draw[->, dashed, thick, black] (fs0.east) -- (fs3.west);
			% Arrows Bottom
			\draw[<->, thick, black] (-11,-7.5) -- (-2.1,-7.5)node[midway, below, font=\footnotesize] {User-Facing Components};
			\draw[<->, thick, black] (-1.9,-7.5) -- (7,-7.5)node[midway, below, font=\footnotesize] {Service Components};
		\end{tikzpicture}
	\end{adjustbox}
	\caption{DESIGN Component Redundancy}\label{service_redundancy}
\end{figure}

@@

This ensures that any change triggered by a node within the DaaS domain
is immediately available to any other node relying in some way
on that particular information change.
Additionally, any extra information needed by a component must be fully contained
within the related request or response.
As shown in figure~\ref{message_routing}, this enables arbitrarily distributed configurations as long as request
and response messages reach their desired target eventually
and as long relevant backend services remain available.

@@

\begin{figure}
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tikzpicture}[every node/.style={minimum width=4cm, minimum height=1cm, align=center}]
			% Frontends
			\node[draw,minimum height=0.5cm, fill=gray!10] (front1) at (-4,0.3) {Frontend 1};
			\node[draw,minimum height=0.5cm, fill=gray!10] (front2) at (0,0.3) {Frontend 2};
			\node[draw,minimum height=0.5cm, fill=gray!10] (front3) at (+4,0.3) {Frontend 3};
			% Backends
			\node[draw,minimum height=0.5cm, fill=gray!30] (back1) at (-4,1.1) {Backend 1};
			\node[draw,minimum height=0.5cm, fill=gray!30] (back2) at (+0,1.1) {Backend 2};
			\node[draw,minimum height=0.5cm, fill=gray!30] (back3) at (+4,1.1) {Backend 3};
			% Instances
			\node[draw,minimum width=12cm, minimum height=1cm, fill=gray!80] (instdomain) at (+0,2.1) {};
			\node[draw,minimum height=0.5cm, fill=gray!80] (inst) at (+0,2) {Instance};
			% Message
			\node[draw,minimum height=0.5cm, fill=gray!50] (message) at (0,-0.75) {Message};
			% Arrows instance
			\draw[->, thick, black] (inst.west) -- ++(-3,0) node[midway, above=-5pt, font=\scriptsize] {Switch to host 1};
			\draw[->, thick, black] (inst.east) -- ++(3,0) node[midway, above=-5pt, font=\scriptsize] {Switch to host 3};
			% Arrows backend
			\draw[->, thick, black] (back1.north) -- ++(0,0.25);
			\draw[->, thick, black] (back2.north) -- ++(0,0.25);
			\draw[->, thick, black] (back3.north) -- ++(0,0.25);
			% Arrows frontend
			\draw[->, thick, black] (front1.north) -- (back1.south);
			% \draw[->, thick, black] (front1.north) -- (back2.south);
			% \draw[->, thick, black] (front1.north) -- (back3.south);
			% \draw[->, thick, black] (front2.north) -- (back1.south);
			\draw[->, thick, black] (front2.north) -- (back2.south);
			% \draw[->, thick, black] (front2.north) -- (back3.south);
			% \draw[->, thick, black] (front3.north) -- (back1.south);
			% \draw[->, thick, black] (front3.north) -- (back2.south);
			\draw[->, thick, black] (front3.north) -- (back3.south);
			% Arrows Message
			\draw[->, thick, black] (message.north) -- (front1.south);
			\draw[->, thick, black] (message.north) -- (front2.south);
			\draw[->, thick, black] (message.north) -- (front3.south);
		\end{tikzpicture}
	\end{adjustbox}
	\caption{Message Oriented Routing.}\label{message_routing}
\end{figure}
